---
title: "The role of adult Noctuidae moths and their food plants in a nocturnal pollen-transport network in Mallorca (Balearic Islands, Spain)"
author: "Joan Díaz Calafat"
date: "24/10/2021"
output: pdf_document
---

# Code reproducibility

To allow a better reproducibility of this script, we recommend to use checkpoint() set at 2021-10-24. This allows to install and use the packages that were available and run at the time of making this script

```{r}
#install.packages("checkpoint")
library(checkpoint)
checkpoint("2021-10-24")
```

# Package and data loading

Load packages:

```{r, warning=FALSE, message=FALSE}
library(bipartite)
```

Load data on moth-plant interactions:

```{r}
pol.data.all <- read.csv("Moth_pollen_list.csv", header = TRUE, encoding = "UTF-8", sep = ";")

# Build an interaction matrix
net.all <- data.frame(unclass(table(pol.data.all$Moth.species, pol.data.all$Pollen.species)))

# Remove species with only one interaction:
# Detect plant (col) and moth (row) species with only one interaction
col.out <- names(colSums(net.all)[which(colSums(net.all) == 1)])
row.out <- names(rowSums(net.all)[which(rowSums(net.all) == 1)])

# Remove those from the main dataset
net.all <- net.all[,-which(colnames(net.all) %in% col.out)]

net.all <- net.all[-which(rownames(net.all) %in% row.out),]

# Remove Unknown Ericacaee, since they would either be A. unedo or E. multiflora and therefore creating a new fake species would hamper our analyses

net.all <- net.all[,-which(colnames(net.all) %in% "Ericaceae")]
```

# Bipartite plots

The adjacency matrix of the moth-plant interactions:

_To properly visualize this plot please use the zoom tool in the RStudio plot viewer. Export at 1267x1000 pixels._

```{r}
visweb(net.all, 
       type="nested",
       square="interaction",
       labsize = 1.8
       )
```

The interaction network:

_To properly visualize this plot please use the zoom tool in the RStudio plot viewer. Export at 1267x1000 pixels._

```{r}
# The plot can be seen properly when visualizing it outside of the RMarkdown window

par(font = 3) # turn all text to italics

plotweb(net.all, 
        method="cca", 
        #Set the color of the row nodes
        col.low="black",
        #Set the color of the column nodes, with A. unedo interactions a different color
        col.interaction = ifelse(colnames(net.all) == "Arbutus.unedo" ,
                             adjustcolor('darkgreen', alpha.f = 0.5), #add transparency to colors
                             adjustcolor('black', alpha.f = 0.5)),
        bor.col.interaction = NA,
        col.high="dark green", #Set the link color
        #Set the rotation of node labels
        text.rot="90", 
        #Set the size of node labels
        labsize=1.8,
        y.lim = c(-0.3, 2.1))
```

# Network analyses

## Metric calculation at the full network level

```{r}
ind.all <- networklevel(net.all, index = c("mean number of shared partners", "partner diversity", "niche overlap", "extinction slope", "robustness", "connectance", "weighted connectance", "web asymmetry", "H2"))

grouplevel(net.all, index=c("mean number of species", "mean number of links"))

ind.all

write.csv2(ind.all, "indices_network.csv") #Export
```

## Metric calculation at the species level

```{r}
ind.all.sp <- specieslevel(net.all, index=c("degree", "species strength", "interaction push pull", "species specificity"), level="both", logbase=exp(1), low.abun=NULL, 
	high.abun=NULL, PDI.normalise=TRUE, PSI.beta=c(1,0), nested.method="NODF", 
	nested.normalised=TRUE, nested.weighted=TRUE, empty.web=TRUE)

ind.all.sp[1]
head(ind.all.sp[2])

write.csv2(ind.all.sp[1], "ind_all_sp_plants.csv") #Export
write.csv2(ind.all.sp[2], "ind_all_sp_moths.csv") #Export
```